\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

This study presents a determinant-based workflow for predicting antimicrobial resistance (AMR) phenotypes in \textit{Salmonella enterica} from curated genomic evidence. Rather than using genome-wide features such as k-mers or pan-genome encodings, the study focuses on biologically interpretable AMR determinants derived from AMRFinderPlus and the NCBI MicroBIGG-E resource. The main objective is to determine how well these curated determinants can predict binary resistant/susceptible phenotypes and how prediction behavior changes when genotype evidence is connected to phenotype under different levels of biological constraint.

The workflow begins by aligning phenotype and genotype records through shared \texttt{BioSample} identifiers. Phenotype processing removes ambiguous, contradictory, and duplicate AST records and retains only resistant and susceptible labels. Genotype processing removes exact duplicates, collapses repeated detections of the same determinant within an isolate, normalizes combined class/subclass annotations, and resolves important edge cases such as multidrug combinations and combination antibiotics. After normalization, antibiotic-specific model-input tables are generated under three determinant-connection scopes: \textit{strict}, \textit{broad}, and \textit{all}. The \textit{strict} scope keeps only narrow lineage-matched determinants, \textit{broad} expands matching to the broader drug class to capture cross-resistance, and \textit{all} uses the entire determinant set observed in the isolate, which maximizes coverage but risks co-occurrence leakage.

Using a shared cleaned cohort, the final experiments evaluate 14 antibiotics that satisfy stable modeling thresholds. One XGBoost classifier is trained per antibiotic per scope with five-fold stratified cross-validation, using resistance as the positive class. The final engineered features retain determinant coverage, determinant identity, and lineage-match information. In parallel, curated rule-based baselines are constructed under the \textit{strict} and \textit{broad} settings to represent direct genotype-to-phenotype inference without learned combination logic.

The results show a clear pattern. The \textit{all} scope provides the strongest overall predictive performance and completely removes the zero-feature problem. The \textit{broad} scope is the strongest biologically constrained alternative and performs especially well when broader within-class resistance mechanisms must be captured, such as beta-lactam cross-resistance. The \textit{strict} scope remains valuable as a narrow biological reference but is too fragile to serve as a universal default because it leaves many isolates with no usable determinant evidence. Feature-importance analysis further shows that top-ranked signals often align with known mechanisms, while wider scopes also reveal the risk of learning co-occurring but not necessarily causal resistance features.

Overall, the study shows that determinant-based AMR prediction is a suitable and interpretable bioinformatics approach for this problem. More importantly, it demonstrates that the genotype-to-phenotype connection rule is itself a major scientific variable: the choice between \textit{strict}, \textit{broad}, and \textit{all} changes not only predictive performance, but also the biological meaning of the resulting model.
