\chapter*{Glossary}
\phantomsection
\addcontentsline{toc}{chapter}{Glossary}

\begingroup
\small
\renewcommand{\arraystretch}{1.02}
\setlength{\LTpre}{0pt}
\setlength{\LTpost}{0pt}
\begin{longtable}{|p{0.24\textwidth}|p{0.69\textwidth}|}
\hline
\textbf{Term} & \textbf{Description} \\ \hline
\endfirsthead
\hline
\textbf{Term} & \textbf{Description} \\ \hline
\endhead
AMR & Antimicrobial resistance. \\ \hline
AST & Antimicrobial susceptibility testing. \\ \hline
Genotype & DNA-based representation of an isolate, summarized here by detected resistance genes and resistance-associated mutations. \\ \hline
Phenotype & Observed resistant or susceptible outcome from laboratory testing. \\ \hline
Determinant & A curated resistance gene or resistance-associated mutation used as biologically meaningful evidence. \\ \hline
MIC & Minimum inhibitory concentration, the lowest drug concentration that prevents visible bacterial growth. \\ \hline
BioSample & The isolate identifier used to match genotype and phenotype records across resources. \\ \hline
Cross-resistance & Resistance mechanism whose effect extends to multiple related antibiotics within the same family or class. \\ \hline
Co-resistance & Co-occurrence of multiple resistance determinants in the same isolate, even when they are not equally relevant to one target antibiotic. \\ \hline
Beta-lactam & A major antibiotic class that includes penicillins, cephalosporins, and related drugs. \\ \hline
Strict scope & Connection scope that keeps only narrowly matched determinant evidence for the target antibiotic. \\ \hline
Broad scope & Connection scope that allows determinant evidence from the broader class of the target antibiotic. \\ \hline
All scope & Connection scope that retains all curated AMR determinants detected in the isolate. \\ \hline
ROC-AUC & Score that measures how well a model separates resistant and susceptible isolates across decision thresholds. \\ \hline
Zero-feature row & Sample whose final input vector contains no retained determinant evidence under a given scope. \\ \hline
\end{longtable}
\endgroup
