\chapter{Background Knowledge}

\section{Antimicrobial Resistance}

% Ý đoạn: định nghĩa AMR và nêu ba con đường hình thành kháng ở mức khái quát:
% đột biến, nhận vật liệu di truyền mới, và áp lực chọn lọc do dùng thuốc.
Antimicrobial resistance (AMR) refers to the ability of bacteria to withstand the action of antimicrobial drugs. At the bacterial level, resistance can arise through mutational adaptations, acquisition of genetic material, or alteration of gene expression. In particular, resistance may occur either by mutations or by the acquisition of resistance-conferring genes through horizontal gene transfer. Over recent decades, the rate at which these processes occur and the number of resistant strains have increased markedly under selective pressure from human antibiotic use~\citep{blair2014mechanisms,munita2016mechanisms,vonwintersdorff2016dissemination}.

% Ý đoạn: giới thiệu các nhóm cơ chế kháng chính ở mức sinh học nền, không đi sâu ví dụ kỹ thuật.
At a broad biological level, AMR can be organized into several major mechanism categories. One category is reduced drug uptake, in which bacteria limit antibiotic entry into the cell. Another is active efflux, in which transport systems extrude the harmful molecule and lower its intracellular concentration. Resistance can also arise through target-site modification, which reduces drug affinity, and through drug inactivation, in which bacteria enzymatically modify or destroy the antimicrobial compound. In addition, some resistant phenotypes involve changes to, or bypassing of, the original target site~\citep{blair2014mechanisms,munita2016mechanisms}.

% Ý đoạn: chốt lại rằng AMR không phải một cơ chế đơn lẻ mà là tập hợp nhiều chiến lược thích nghi.
These mechanisms show that AMR is not a single biological event but a collection of adaptive strategies. Different antibiotics may be affected by different mechanisms, and the same bacterial population may accumulate multiple resistance traits through mutation, altered expression, and horizontally acquired determinants. For that reason, understanding AMR requires both a general view of resistance biology and a more specific understanding of how particular genomic determinants relate to antimicrobial families~\citep{blair2014mechanisms,munita2016mechanisms,vonwintersdorff2016dissemination}.


\section{Genotype and Phenotype}

\subsection{Genotype}

% Ý đoạn: định nghĩa genotype trong bối cảnh AMR và giới thiệu AMR determinants
% như phần thông tin di truyền có liên quan đến kháng thuốc.
In AMR studies, genotype refers to the bacterial genomic information that may encode, support, or influence antimicrobial resistance. In this context, the most relevant genomic features are AMR determinants, meaning genes or mutations that provide biologically meaningful resistance-related evidence~\citep{munita2016mechanisms,feldgarden2021amrfinderplus}.

% Ý đoạn: làm rõ determinants không chỉ là gene mà còn gồm point mutations,
% đồng thời phân biệt acquired determinants với chromosomal determinants.
AMR determinants are broader than resistance genes alone. They include not only acquired resistance genes but also resistance-associated point mutations. This distinction is important because resistance may arise either through the acquisition of a new resistance factor or through mutational changes that alter an endogenous target or pathway. It is therefore also useful to distinguish acquired determinants from chromosomal determinants. Acquired determinants are commonly introduced through horizontal gene transfer, often on mobile genetic elements such as plasmids or transposons. Chromosomal determinants, by contrast, usually correspond to endogenous genomic changes within the bacterial genome~\citep{blair2014mechanisms,munita2016mechanisms,vonwintersdorff2016dissemination,feldgarden2021amrfinderplus}.

% Ý đoạn: giải thích vì sao determinants cần được hiểu theo ngữ cảnh class/subclass
% của kháng sinh, để nối sang phần genotype-phenotype relation.
AMR determinants are also interpreted in relation to antibiotic classes and subclasses. A determinant may be associated with one specific drug, with a subgroup of related drugs, or with a broader antimicrobial family. This means that genotype in AMR research should not be understood as a collection of isolated genes or mutations without context. Instead, it should be interpreted as a structured representation of resistance-related genomic evidence whose meaning depends on determinant type, biological mechanism, and antibiotic context~\citep{feldgarden2021amrfinderplus}.

\subsection{Phenotype}

% Ý đoạn: định nghĩa phenotype trong bối cảnh AMR và phân biệt với genotype.
Phenotype refers to the observable response of a bacterial isolate when it is exposed to an antimicrobial agent. In the AMR setting, phenotype is concerned with whether growth is inhibited under defined laboratory conditions. Whereas genotype describes resistance-related genomic potential, phenotype captures the measured drug-response outcome observed in testing~\citep{shen2024comparison}.

% Ý đoạn: giới thiệu AST là cách đo phenotype và MIC là đại lượng đo cốt lõi.
The main laboratory approach for characterizing AMR phenotype is antimicrobial susceptibility testing (AST). A central concept in AST is the minimum inhibitory concentration (MIC), which is the lowest concentration of an antibacterial agent that prevents visible growth under strictly controlled in vitro conditions. MIC is therefore a quantitative laboratory measurement of phenotypic response rather than an interpretation by itself~\citep{kowalskakrochmal2021mic,doern2011ast}.

% Ý đoạn: giải thích breakpoint dùng để chuyển MIC thành các nhãn phenotype như S/I/R.
To support clinical and surveillance use, MIC values are interpreted through breakpoint criteria and translated into categories such as susceptible, intermediate, and resistant. In other words, breakpoints serve to convert numerical AST results into clinically interpretable phenotype labels. In many AMR studies, these categorical outputs are used as the phenotype targets for downstream analysis~\citep{doern2011ast,jenkins2011breakpoints,shen2024comparison}.

\subsection{Relationship Between Genotype and Phenotype}

% Ý đoạn: nêu ý chính rằng genotype cung cấp bằng chứng kháng thuốc,
% nhưng không ánh xạ hoàn hảo sang phenotype quan sát được.
Although genotype provides biologically meaningful evidence about resistance potential, it does not map perfectly to phenotype. The presence of a known determinant does not always guarantee that laboratory resistance will be observed, and the absence of an identified determinant does not always imply susceptibility. This mismatch is one of the central biological challenges in AMR research.

% Ý đoạn: tóm tắt các nguyên nhân chung làm cho ánh xạ genotype-phenotype
% trở nên không hoàn hảo.
Several factors contribute to this difficulty. Biological knowledge remains incomplete, so some resistance mechanisms may still be unknown or only partially characterized. In addition, gene expression, regulatory effects, and variant-specific behavior can influence whether a determinant produces a measurable resistant phenotype. Phenotype may also depend on interactions among multiple determinants, as well as on environmental conditions, testing procedures, and breakpoint interpretation.

% Ý đoạn: giới thiệu cross-resistance và co-resistance như hai nguồn làm
% quan hệ genotype-phenotype trở nên phức tạp hơn.
Two related concepts further complicate this relationship. Cross-resistance occurs when one biological mechanism affects multiple related antibiotics, so the relevance of a determinant may extend beyond a single drug label~\citep{munita2016mechanisms}. Co-resistance occurs when multiple determinants are present in the same isolate, often because resistance genes can spread together on mobile genetic elements or be maintained under shared selection pressures~\citep{vonwintersdorff2016dissemination}. For these reasons, linking genomic evidence to observed antimicrobial response is biologically informative but not straightforward.
