\section{Research Objectives}

This study is guided by the following objectives:

\begin{enumerate}[label=\arabic*.]
    \item Build a reproducible determinant-based AMR prediction workflow that connects curated genotype data, AST phenotype labels, and antibiotic grouping knowledge through a shared isolate identifier.
    \item Clean phenotype data so that the final labels are restricted to a stable binary resistant/susceptible formulation suitable for supervised learning.
    \item Clean and normalize genotype data so that duplicated detections, combined class/subclass labels, and known edge cases do not distort downstream antibiotic-specific feature construction.
    \item Construct three determinant-connection settings---\textit{strict}, \textit{broad}, and \textit{all}---to study how biological scope changes the usable feature space.
    \item Train and evaluate antibiotic-specific machine-learning models under a consistent cross-validation protocol using resistance as the positive class.
    \item Include curated determinant-based rule baselines under constrained scopes so that direct genotype-to-phenotype inference can be compared against learned models.
    \item Compare the resulting methods primarily in terms of resistant-case recall, resistant-case precision, and ROC-AUC.
    \item Interpret the observed differences biologically, especially with respect to narrow lineage matching, cross-resistance within drug classes, zero-feature rows, and possible co-occurrence leakage.
\end{enumerate}

Taken together, these objectives define two main deliverables. First, the study provides a detailed and reproducible methodology for building determinant-based AMR prediction datasets from surveillance-style resources. Second, it provides an empirical comparison showing when narrower biologically constrained inputs are sufficient, when broader class-level evidence is beneficial, and when the widest determinant scope improves prediction at the cost of interpretability risk.

The remainder of the report is organized as follows. Chapter 2 introduces the biological background relevant to AMR, genotype, and phenotype, Chapter 3 describes the workflow and modeling design, and Chapter 4 presents the experimental results and limitations. The report closes with a conclusion summarizing the main findings.
