\section{Problem Definition}

% Ý đoạn: định nghĩa bài toán AMR prediction theo nghĩa khái niệm,
% rồi giới thiệu Salmonella như case study cụ thể.
% Vai trò: đặt phạm vi của bài toán trước khi đi vào formulation chi tiết hơn.
Antimicrobial resistance prediction from genomic data is the task of estimating a laboratory resistance phenotype from DNA-derived evidence. More generally, it is a genotype-to-phenotype inference problem in which resistance-related genomic information is used to anticipate observed antimicrobial response~\citep{shen2024comparison,ren2022prediction}. The scientific interest of this study is therefore broader than one organism or one dataset. In this report, \textit{Salmonella enterica} is used as the case study because it is well represented in surveillance resources and provides enough paired genotype and phenotype records for a credible antibiotic-specific evaluation setting.

% Ý đoạn: mô tả input-output của bài toán theo mức vừa đủ,
% nhưng chưa đi sâu vào workflow hay implementation.
% Vai trò: giúp người đọc hình dung rõ bài toán dự đoán mà không biến section này thành methodology.
At the isolate level, the input is resistance-related genomic evidence and the output is an antibiotic-specific phenotype derived from antimicrobial susceptibility testing (AST). 
In the present study, this genomic evidence is represented through curated AMR determinants, including acquired genes and chromosomal mutations.
~\citep{feldgarden2021amrfinderplus,munita2016mechanisms}. 
The practical question is therefore whether the determinant profile of a single isolate can reliably predict the final AST phenotype for a target antibiotic.


% Ý đoạn: nêu khó khăn cốt lõi là khoảng cách giữa genotype và phenotype.
% Vai trò: làm rõ vì sao đây không chỉ là bài toán huấn luyện classifier thông thường.
The difficulty of the problem lies in the gap between genotype and phenotype. Some resistance mechanisms are directly associated with one antibiotic, whereas others act more broadly across a drug class. In addition, a detected determinant may co-occur with other resistance signals without being the direct cause of the target phenotype~\citep{shen2024comparison,munita2016mechanisms}. As a result, the main challenge is not only to train a classifier, but also to define a biologically meaningful representation of genotype evidence before prediction.

% Ý đoạn: rút toàn bộ phần problem definition về một research question trung tâm.
% Vai trò: tạo trục chính cho toàn report và cho các phần sau bám vào.
Accordingly, this study asks whether curated AMR determinants can reliably predict antibiotic resistance phenotype, and how prediction quality changes when genotype evidence is connected to phenotype under different levels of biological constraint.

% Ý đoạn: chốt lại study position của đề tài.
% Vai trò: nhấn mạnh đây là nghiên cứu về ảnh hưởng của biological constraint,
% không chỉ là một benchmark mô hình dự đoán.
The study examines how curated genomic evidence supports antibiotic-specific AMR prediction under different levels of biological constraint while remaining biologically interpretable. The \textit{Salmonella} cohort provides a concrete experimental setting with sufficient paired genotype and phenotype records, but the underlying question is broader than a single organism. At the center of the report is how biologically constrained determinant evidence changes antibiotic-specific prediction behavior.
