\section{Related Work}

% Ý đoạn: mở phần related work bằng bài toán chung và tension chính trong literature:
% predictive coverage so với biological interpretability.
% Vai trò: cho người đọc biết vì sao cần phân loại literature thay vì chỉ liệt kê paper.
Genomic antimicrobial resistance prediction is already an active research direction, and recent reviews show that the field has expanded rapidly across different pathogens, data representations, and machine-learning strategies~\citep{kim2022ml,tang2022systematic,ardila2024wgsml}. The central goal is to infer observed resistance phenotype from genomic evidence quickly and at scale, but this goal remains difficult because genotype does not map perfectly to phenotype. Known resistance genes and mutations provide useful biological clues, yet they do not always explain the final laboratory phenotype completely. At the same time, broader genome-wide representations may improve predictive coverage but often weaken direct biological interpretation.

% Ý đoạn: giới thiệu logic tổ chức literature thành nhiều hướng tiếp cận.
% Vai trò: chuẩn bị cho ba family bên dưới và nêu rõ tiêu chí so sánh của report.
As a result, the literature has not converged on one universally best solution. Some studies emphasize curated biological knowledge and direct interpretability, some use biologically meaningful features together with machine learning, and others move toward broad genome-wide representations such as SNPs, k-mers, or pan-genome features~\citep{feldgarden2021amrfinderplus,maguire2019identification,ren2022prediction}. Recent reviews also highlight that these choices affect not only predictive performance but also interpretability, robustness, and potential clinical utility~\citep{kim2022ml,tang2022systematic,ardila2024wgsml}. This diversity is important for the present report because the main scientific question is not only whether AMR can be predicted, but also how the choice of genotype-to-phenotype representation affects the balance between predictive performance and biological explanation.

% Ý đoạn: giải thích chức năng của bảng related work.
% Vai trò: nói rõ bảng này dùng để định vị phương pháp, không phải benchmark số học trực tiếp.
Table~\ref{tab:related-work-overview} summarizes representative studies that are most relevant to the present project across organism, phenotype target, feature representation, and evaluation design.

% Ý đoạn: family 1 = rule/knowledge-driven, mạnh về interpretability nhưng hạn chế về coverage.
% Vai trò: tạo mốc so sánh cho hướng baseline sinh học trực tiếp của đề tài.
The first family of studies begins from known AMR genes, known resistance-associated mutations, and curated biological annotation. In this direction, the main advantage is interpretability: every prediction or association can be traced back to a specific determinant or mechanism. The AMRFinderPlus framework of Feldgarden et al.~\citep{feldgarden2021amrfinderplus} is especially important here because it provides the curated determinant catalog that motivates the feature representation used in this report. Shen et al.~\citep{shen2024comparison} is also important because it directly compares genotypic and phenotypic resistance profiles in \textit{Salmonella enterica}, which is close in spirit to the rule-based reference baselines used in the present study. The limitation of this family is that direct inference can fail when resistance depends on broader within-class mechanisms, unknown variants, or incomplete determinant coverage.

% Ý đoạn: family 2 = biology-informed machine learning.
% Vai trò: xác định nhóm tài liệu gần nhất với hướng chính của đề tài.
The second family uses biologically meaningful features but lets machine learning learn the final mapping from genotype to phenotype. This is the closest family to the present project. Maguire et al.~\citep{maguire2019identification} is particularly relevant because it combines known AMR gene annotations with supervised learning and also contrasts them with gene-free k-mer-based models in \textit{Salmonella}. In this family, the main advantage is that the model can capture determinant interactions and weighting effects that a fixed rule cannot express easily. He et al.~\citep{he2025prediction} extends this direction toward broader pan-genome and pan-resistome modeling in \textit{Salmonella}, which is particularly relevant when comparing our biologically constrained settings with the wider \textit{all}-scope representation.

% Ý đoạn: family 3 = genome-wide / ML-heavy.
% Vai trò: nêu hướng coverage cao hơn nhưng đánh đổi interpretability, để làm đối trọng cho đề tài.
The third family moves toward broader genome-wide representations and more feature-heavy machine-learning designs. Ren et al.~\citep{ren2022prediction} predicts antimicrobial resistance from whole-genome sequencing using SNP-based features, whereas Wang et al.~\citep{wang2023using} models MIC values for nontyphoidal \textit{Salmonella} from genome-derived 10-mers. These studies illustrate that predictive signal can extend beyond curated determinant catalogs and that broad genomic representations can perform strongly. At the same time, their feature spaces are less directly interpretable biologically. For the present report, they are therefore most useful as high-coverage comparison directions rather than as directly comparable determinant-based baselines.

% Ý đoạn: phát biểu research gap một cách trực tiếp.
% Vai trò: nói rõ literature còn thiếu gì và tại sao study này cần thiết.
Taken together, prior work has explored rule-based inference, determinant-based machine learning, and broader genome-wide prediction. Within the literature reviewed here, however, genotype-to-phenotype connection scope does not appear to have been treated explicitly as the main experimental variable in a determinant-based AMR prediction workflow. In other words, existing studies show that genomic AMR prediction is feasible, but they do not directly focus on how prediction behavior changes when curated determinant evidence is connected to phenotype under narrower or broader biological scope assumptions.

% Ý đoạn: self-positioning của nghiên cứu hiện tại trong landscape của literature.
% Vai trò: cho biết đề tài đứng ở đâu và khác gì so với các hướng trước đó.
The study is most closely aligned with the biology-informed machine-learning family, while remaining situated between strict rule-based inference and broader genome-wide prediction. Curated determinant-level features preserve interpretability, rule-based baselines provide constrained references, and multiple connection scopes make it possible to compare how biological restriction changes prediction behavior under a matched antibiotic-specific setting.

% Ý đoạn: chốt contribution ở mức conceptual.
% Vai trò: nhấn mạnh report này là controlled comparison về biological scope,
% không phải chỉ thêm một classifier benchmark.
Within this landscape, determinant-based prediction occupies an intermediate position between direct deterministic matching and broader black-box genomic representations. The resulting analysis is therefore a controlled comparison of biological connection scopes in genomic AMR prediction.

\begin{table}[]
\centering
\caption{Representative related studies and their relation to the present project}
\label{tab:related-work-overview}
\scriptsize
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{@{}p{2.1cm} p{1.8cm} p{2.1cm} p{2.0cm} p{4.8cm}@{}}
\toprule
Study & Target & Feature type & Method family & Relation to our study \\
\midrule
Feldgarden et al.~\citep{feldgarden2021amrfinderplus} & General AMR annotation & Curated determinants & Biology-first resource & Foundation for the determinant space used in our pipeline; not a direct prediction benchmark. \\
Shen et al.~\citep{shen2024comparison} & Salmonella concordance & Genotypic AMR profiles & Rule-based direction & Closest conceptual basis for our direct genotype-to-phenotype baseline. \\
Maguire et al.~\citep{maguire2019identification} & Salmonella AMR prediction & AMR gene annotations / k-mers & Biology-informed ML & Shows both determinant-centered and gene-free learning within the same Salmonella task. \\
He et al.~\citep{he2025prediction} & Salmonella phenotype and MIC & Pan-genome / pan-resistome & Broader ML & Closest literature analogue to our wider-scope machine-learning direction. \\
Ren et al.~\citep{ren2022prediction} & Binary AMR prediction & Whole-genome SNPs & Genome-wide ML & Useful for discussing broad genomic signal beyond curated determinants. \\
Wang et al.~\citep{wang2023using} & Salmonella MIC prediction & WGS-derived 10-mers & Genome-wide ML & Useful mainly for the richer MIC target and the interpretability trade-off of genome-wide features. \\
\bottomrule
\end{tabular}
\end{table}
